Bayesian exponential random graph modeling of KWL Burial network

### 2020-10-23 analysis/paper/paper.Rmd
add code for burial value, run model, and edit Intro

<!--burial study- network analysis-->
Burials contains valuable information about individual mortuary practices and inherited structure of a society as a whole. With the development of network science and graph theory, it is promising to study relation among burials and explore its social implication [@Sosna2013]. Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals, allowing the inference for social structures. This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom. Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp. 252]. Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization. A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017]. Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007]. Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes. The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016]. In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relations in societies [@Marin2011]. With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affects the structure [@Freeman2004; @Salvini2010]. The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]

<!--archaeology case-->
The archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and accumulation of them in residential area around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in the society. Those prestige goods include imported ornaments and porcelains that entered northeastern Taiwan through regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups. In addition to prestige, European colonial powers could have provided military support to Indigenous allies. For example, @Voss2005's study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local Indigenous people. A general model to summarize these case studies might be that the support of a colonial power combined with high local values of imported goods, might lead to increased social inquqality due to competition among individuals. In this study, we use this pericolonial model to explain the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century. Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and test our hypothesis that increased inequality observed in social network after the European arrival. This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.

We applied Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan. We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group. We test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions. We ask the questions: 1. Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities? 2. Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an Indigenous society can be detected by burials networks. 3. How the relations among burials may affect the quantity of burial goods? We expect the network to present a higher degree of clustering (e.g. the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period. This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups. A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.

@@

We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan. We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here. We treat the presence of relations between burials based on number or types of  prestige items from regional trade network. where a few prestige goods controlled by a small number of individuals. This is because prestige items could represent a relationship, where a few prestige goods controlled by a small number of individuals. Based on this assumption, we treat the presence of relations between burials through the burial goods in comoon.  Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we classifed those beads into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index. Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

>>>>

<!--burial study- network analysis-->
Burials contains valuable information of both individual mortuary practices and the structure of a society inherited in burial relations. With the development of network science and graph theory, studying relations among burials and exploring its social implication become promising [@Sosna2013]. Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals, allowing the inference for social structures. This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom. Burials with grave goods in common often indicate some underlying social relations connected through trade, exchange, and gifting [@Coward2013, pp. 252]. Identifying the structure of relationships among burials can give an insight to the structure of a social organization. One common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies in recent years [@Mills2017]. Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007]. Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes. The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016]. In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, indicating social relations in societies [@Marin2011]. With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010]. The application in archaeology is widely used for different topics including distribution of archaeological sites, grave goods, and social complexity [@Coward2013; @Mizoguchi2013; @Potgieter2008]

<!--archaeology case-->
The archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and some accumulation patterns in residential area around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in those societies. The prestige goods include imported ornaments and porcelains that entered northeastern Taiwan via regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups. The concept of prestige goods may develop earlier before the European arrival, but could be amplified during the European colonial period. For example, a person with more imported goods may be treated as higher status. A general model to summarize these case studies might be that the support of a colonial power combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals. In this study, we use a pericolonial model to explain the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century. Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and test our hypothesis that increased inequality observed in social network after the European arrival. This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.

We applied Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan. We viewed individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same burial goods that can represent the same social group with similar economic or political status. We construct a network that represents higher social inequality and tested hypotheses that burial relations becomes more hierarchical after the European arrival. We ask the questions: 1. Do the observed burial data after the European presence shows more similar network quality as our hypothesized model? 2. Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an Indigenous society can be detected by burials networks. 3. What is the major variables affecting or forming the higher degree of popularity (e.g. a few node has more relationship with other nodes) in the European contact period that might hint at social heterogeneity during this period. This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups. A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.

@@

We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan. We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here. We treat the presence of relations between burials based on number or types of  prestige items from regional trade network. where a few prestige goods controlled by a small number of individuals. This is because prestige items could represent a relationship, where a few prestige goods controlled by a small number of individuals. Based on this assumption, we treat the presence of relations between burials through the burial goods in common.  Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we classified those beads into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index. Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
